\documentclass[10pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ mathrsfs }
\usepackage{ mathtools}
\usepackage{enumerate}
\usepackage{bbm}
\usepackage{lipsum}
\usepackage{fancyhdr}
\usepackage{tikz-cd} 
\usepackage{adjustbox} 
\numberwithin{equation}{section}
\usetikzlibrary{arrows}
\newcommand{\midarrow}{\tikz \draw[-triangle 90] (0,0) -- +(.1,0);}
\tikzset{commutative diagrams/.cd,
mysymbol/.style={start anchor=center,end anchor=center,draw=none}
}
\newcommand\comsymb[2][\circlearrowleft]{%
  \arrow[mysymbol]{#2}[description]{#1}}

\newtheorem{theorem}{Theorem} [section] 
\newtheorem{proposition}{Proposition}[section] 
\newtheorem{definition}{Definition}[section] 
\newtheorem{lemma}{Lemma}[section] 
\newtheorem{notation}{Notation}[section] 
\newtheorem{remark}{Remark}[section] 
\newtheorem{corollary}{Corollary} [section] 
\newtheorem{terminology}{Terminology}[section] 
\newtheorem{fact}{Fact}[section] 
\newtheorem{example}{Example}[section] 
\newtheorem{claim}{Claim}
\newenvironment{claimproof}[1]{\par\noindent\underline{Proof:}\space#1}{\hfill $\blacksquare$}

\DeclareMathOperator{\tmod}{mod}
\DeclareMathOperator{\triv}{triv}
\DeclareMathOperator{\ind}{ind}
\DeclareMathOperator{\straw}{s.t.}
\DeclareMathOperator{\ev}{ev}
\DeclareMathOperator{\pr}{pr}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Map}{Map}
\DeclareMathOperator{\Stab}{Stab}
\DeclareMathOperator{\Ind}{Ind}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\SO}{SO}
\DeclareMathOperator{\Sp}{Sp}
\DeclareMathOperator{\esssup}{esssup}
\DeclareMathOperator{\abs}{abs}
\DeclareMathOperator{\rel}{rel}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Homk}{Hom_{k-alg}}
\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\Image}{Im}
\DeclareMathOperator{\Dom}{Dom}
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\divergence}{div}
\DeclareMathOperator{\rk}{rk}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\mSpec}{mSpec}
\DeclareMathOperator{\MaxSpec}{MaxSpec}
\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\Mat}{Mat}
\DeclareMathOperator{\PD}{PD}
\DeclareMathOperator{\ext}{ext}
\DeclareMathOperator{\vol}{vol}
\DeclareMathOperator{\inte}{int}
\DeclareMathOperator{\diverg}{div}
\DeclareMathOperator{\curl}{curl}
\DeclareMathOperator{\image}{im}
\DeclareMathOperator{\dR}{dR}
\DeclareMathOperator{\Ob}{Ob}
\DeclareMathOperator{\Mnfd}{Mnfd}
\DeclareMathOperator{\OpEmb}{OpEmb}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Spa}{Spa}
\DeclareMathOperator{\Lim}{Lim}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\rid}{rid}
\DeclareMathOperator{\Spf}{Spf}
\DeclareMathOperator{\Rig}{Rig}
\DeclareMathOperator{\Adic}{Adic}
\DeclareMathOperator{\FSch}{FSch}
%\newcommand*{\name}[\num_arguments][default values]{{\color{#1}\Large #2}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\Ouv}{\mathcal{O}}
\newcommand{\norm}[1]{\Vert #1 \Vert}
\newcommand{\opNorm}[2]{\norm{#1}_{#2\to#2}}
\newcommand{\normL}[3]{\norm{#1}_{L^{#2}(#3)}}
\newcommand{\N}[0]{\mathbb{N}}
\newcommand{\m}[0]{\mathfrak{m}}
\newcommand{\R}[0]{\mathbb{R}}
\newcommand{\Z}[0]{\mathbb{Z}}
\newcommand{\C}[0]{\mathbb{C}}
\newcommand{\Q}[0]{\mathbb{Q}}
\newcommand{\F}[0]{\mathbb{F}}
\newcommand{\G}[0]{\mathbb{G}}
\newcommand{\A}[0]{\mathbb{A}}
\newcommand{\T}[0]{\mathbb{T}}
\newcommand{\Para}[0]{\mathbb{P}}
\newcommand{\M}[0]{\mathcal{M}}
\newcommand{\maniN}[0]{\mathcal{N}}
\newcommand{\cinf}[0]{\mathcal{C}^\infty}
\newcommand{\torus}[0]{\mathbb{T}}
\newcommand{\sep}[0]{\:|\:}
\newcommand{\pd}[2]{{\frac {\partial #1} {\partial #2}}}
\newcommand{\compinc}[0]{\subset \subset}
\newcommand{\sH}[0]{\mathscr{H}}
\newcommand{\ab}[1]{\vert #1 \vert}
\newcommand{\st}[0]{\:\straw\:}
\newcommand{\fib}[1]{%
  \mathbin{\mathop{\times}\limits_{#1}}%
}
\newcommand{\tens}[1]{%
  \mathbin{\mathop{\otimes}\displaylimits_{#1}}%
}

\title{Rigid Analytic Geometry Tutorial}
\author{So Murata}
\date{SoSe 26, University of Bonn}


\begin{document}
\maketitle

%4/20

\section{Week 2}
\subsection{Motivation and Obstacles}
The motivation for the course is that, given a non-archimedian field $k$, how should we define analytic varieties and presheaves of algebraic functions over $k$.

\begin{definition}[Classical Rigid Analytic Geometry]
  Intuition, over $\overline{B(0,1)}^n\subseteq K^n$, the analytic functions should be of the form of convergent power series.
  \begin{equation*}
    f = \sum_{\alpha\in\N^n}a_\alpha x^\alpha.
  \end{equation*}
  This is a local description of the functions. We denote the set of such functions 
  \begin{equation*}
    \T_n(K)\defeq K\langle x_1,x_n\rangle,
  \end{equation*}
  which we call the Tate algebra. 
\end{definition}

\begin{remark}
  There is a one-to-one correspondence between 
  \begin{equation*}
    \Sp\T_n(K)\longleftrightarrow \overline{B_{\overline{K}}(0,1)}^n/\Gal(\overline{K}/K).
  \end{equation*}
\end{remark}

We will denote $X=\Sp\T_n(K)$. Furthermore, we can define admissible opens $U\subseteq X$, analytic functions on $U$, and the presheaf of analytic functions $\Ouv_{X}$.
Over a wide range of admissible open coverings, $\Ouv_{X}$ satisfies the sheaf requirement.

\begin{equation*}
  0\to\Ouv(X)\hookrightarrow\prod_{i}\Ouv(U_i)\rightrightarrows\prod_{i,j}\Ouv(U_i\cap U_j).
\end{equation*}

We only allow certain open coverings called admissible covers. We introduce and work on a Grothendieck topology.
\begin{remark}
  There's a non-trivial sheaf of abelian groups $\mathcal{F}$ on $X$ with $G$-topology such that $\forall x\in X,\F_x=0$.
\end{remark}

To overcome the obstacle above, we will introduce extra points, namely van der Put points.
We glue $\Sp\T_n(K)/I$ along opens to get rigid $K$-spaces.


\subsection{Raynaud's Formal Schemes}

Suppose $X$ is a $\Q_p$-scheme, we want to consider the reduction $\tmod{p}$. We find a $\Z_p$-scheme $X'$ such that 
\begin{equation*}
  X'\fib{\Z_p}\Q_p\cong X.
\end{equation*}
Note that we can't guarantee that $X'$ is unique or even it exists. Take $R$ to be the valuation ring of $K$ for example $K=\Q_p,R=\Z_p$, we want to find an object the structure over $R$, such that its generic fiber is $X$.
Say $X=\Sp\T_n(K)=\Sp K\langle x_1,\cdots,x_n\rangle$, we can take 
\begin{equation*}
  Y = \Spf R\langle x_1,\cdots,x_n\rangle,
\end{equation*}
We also set,
\begin{equation*}
  Y^{\rid}=\Sp(R\langle x_1,\cdots,x_n\rangle\hat{\tens{R}}K).
\end{equation*}
If $R$ is separated $\mathfrak{a}$-adically topologized, that is for any subset $I\subseteq R$, $U$ is open if and only if there is $n\in\N$ such that $\mathfrak{a}^n\subseteq I$, and $\bigcup_{n\geq1}\mathfrak{a}^n=0$. 
Then we have,
\begin{equation*}
  R\langle x_1,\cdots,x_n\rangle = \varprojlim_{m\geq1}R/\mathfrak{a}^n[x_1,\cdots,x_n].
\end{equation*}
We have another construction of rigid spaces, namely Berkovich's theory.
\subsection{Huber's adic spaces}

Let $k$ be algebraically closed and $K/k$ is a finitely generated field extension of transcendence degree $1$. Then there is a non-singular projective curve over $k$ with 
function field $K$. 
\par\textbf{Construction}
Take $C_K$ to be the set of discrete valuation rings in $K/k$. Take closed subsets to be the whole space and finite sets.
On each open $U$, define 
\begin{equation*}
  \Ouv(U) = \bigcap_{p\in U}R_p,
\end{equation*}
then $(C_K,\Ouv)$ is isomorphic to a non-singular curve with function field $K$.
\par Given a topological ring $A$ and and its subring $A^\triangleleft$ or sometimes denoted $A^+$(integrally closed, open, power bounded). Huber defines,
\begin{equation*}
  \Spa(A,A^\triangleleft) = \{\text{continuous valuations $v:A\to \Gamma\cup\{0\}$ such that $\forall a\in A^\triangleleft, v(a)\leq1$}\}/\sim.
\end{equation*}
\begin{remark}  There is a fully-faithful functor 
  \begin{equation*}
    r_K:\Rig(K)\to \Adic/\Spa(K,K^\circ),
  \end{equation*}
  from the category of rigid $K$-spaces, to the category of adic spaces over $\Spa(K,K^\circ)$.
\end{remark}
\begin{remark}  There is a fully-faithful functor 
  \begin{equation*}
    t:\FSch^{\text{ln}}\to \Adic/\Spa(K,K^\circ),
  \end{equation*}
  from the category of locally Noetherian formal schemes, to the category of adic spaces over $\Spa(K,K^\circ)$.
\end{remark}
\begin{remark}
  \label{remark_2_classical_modern_intersection}
  If $X$ is a qcqs rigid $K$-space, the set of van der Put points of $X$ is one-to-one correspondent to points of $r_K(X)$.
\end{remark}
The problem session goes parallel to the lecture, meaning the lecture will focus on the modern treatment while the tutorial will focus on the classical one. Remark \ref{remark_2_classical_modern_intersection} will be the meeting point of these two.
\end{document}